Toward Resolved Simulations of Burning Fronts in Thermonuclear X-ray Bursts

### 2018-09-26 astronum_2018/astronum-xrb.tex
more plot work

We setup a problem by creating a pair of 1D hydrostatic XRB models
representing the post- and pre-flame states and apply these laterally
across the domain to setup a hot region at the left of our domain in
equilibrium with the cool model.  Thermal diffusion and nuclear
reactions quickly setup a laterally propagating flame that begins to
move through our domain.  We resolve the diffusion scale, using a
resolution of 10~cm, and use adaptive mesh refinement (two jumps of
$4\times$) to keep a buffer of low density material at low resolution
above the surface to allow for expansion.  To make the lengthscales
tractable, we rotate the neutron star at 2000~Hz---much faster than
expected.  This helps confine the spreading region to a more
reasonable horizontal region.  Nevertheless, our simulation uses 12288
zones laterally, on the finest grid.  We use the stellar
conducitivities from \cite{Timmes00} and a 13-isotopic alpha-chain
reaction network.  To make the timescales easier, we boost the flame
speed by a factor of 10, by boosting both the conductivity and energy
generation rate.

Figure~\ref{fig:xrb} shows the flame structure.  The top panel shows
the temperature.  The initial perturbation reached to $2\times
10^4~\mathrm{cm}$, so at this point, the extend of the burnt region
has nearly doubled.  The flame structure is quite wide

>>>>

We setup a problem by creating a pair of 1D hydrostatic XRB models
representing the post- and pre-flame states and apply these laterally
across the domain to setup a hot region at the left end of our domain
in equilibrium with the cool model.  Thermal diffusion and nuclear
reactions quickly setup a laterally propagating flame that begins to
move through our domain.  We resolve the diffusion scale, using a
resolution of 10~cm, and use adaptive mesh refinement (two jumps of
$4\times$) to keep a buffer of low density material at low resolution
above the surface to allow for expansion.  To make the lengthscales
tractable, we rotate the neutron star at 2000~Hz---much faster than
expected.  This helps confine the spreading region to a more
reasonable horizontal region.  Nevertheless, our simulation uses 12288
zones laterally, on the finest grid.  We use the stellar
conducitivities from \cite{Timmes00} and a 13-isotopic alpha-chain
reaction network.  To make the timescales easier, we boost the flame
speed by a factor of 10, by boosting both the conductivity and energy
generation rate.  This is an approximation we hope to relax in the
near future.

Figure~\ref{fig:xrb} shows the flame structure for an initial 2D run.
The top panel shows the temperature.  The initial perturbation reached
to $2\times 10^4~\mathrm{cm}$, so at this point, the extent of the
burnt region has nearly doubled.  The flame structure is quite wide,
since there are multiple burning stages beyond just helium to carbon, and
this shows up in the composition.
